---
title: "Lab Week 5"
author: "Allison Horst"
date: "October 26, 2018"
output:
  html_document: default
  word_document: default
  pdf_document: default
subtitle: Graphs and tables
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###0. Get packages and data

In this lab, we'll use the following packages:

- tidyverse
- knitr
- kableExtra
- plotly
- ggrepel
- extrafont

Recall: alt + command + i for inserting code chunk shortcut

```{r, include = FALSE}

library(tidyverse)
library(pwr)
library(knitr)
library(kableExtra)
library(plotly)
library(extrafont)
library(ggrepel)

mortality <- read_csv("drug_mortality.csv")
income <- read_csv("state_income.csv")

```


Today in lab, we'll be: 

- Power analysis
- Wrangling
- Piping directly to ggplot 
- Creating finalized graphics
- Interactive plots with plotly
- Joining data frames (dplyr)
- A nice table (kable + kableExtra)


### 1. Pipe an F-test and a t-test, with in-line references

Check out the built-in dataset chickwts (?chickwts, View(chickwts))
We want to know, are chicks fed 'horsebean' significantly smaller than those fed 'linseed'?

Remember: if doing this for real, we'd look at histograms, qq-plots, test for equal variances (f-test)

```{r}

chick_ftest <- chickwts %>% 
  filter(feed == "horsebean" | feed == "linseed") %>% 
  var.test(weight ~ feed, data = .)

chick_ttest <- chickwts %>%
  filter(feed == "horsebean" | feed == "linseed") %>% 
  t.test(weight ~ feed, data = ., alternative = "less") # (check levels(chickwts))

```

Inline summary of results: 

Weights for chicks fed horsebean are significantly lower than those fed linseed (t(`r round(chick_ttest$parameter,2)`) = `r round(chick_ttest$statistic,2)`, *p* = `r round(chick_ttest$p.value, digits = 3)`, $\alpha$ = 0.05).



### 2. Power Analysis (package 'pwr')

The type of power analysis that you'll do depends on the TEST. Which means that there's a different power calculation for each type of hypothesis test you'll run. 

For example, to find values associated with power for a t-test, you'll use pwr.t.test(), but if you're doing a calculation for an ANOVA, you'll use pwr.anova.test().

Remember, the POWER of a test is the probability that you will detect a significant result if there really is one. It is the complement of committing a Type II Error, $\beta$ (there IS a significant result but you don't detect it). 

First, check out the pwr.t.test() function (make sure 'pwr' package is loaded - if at home, they'll need to install first)

**Example:** You need to collect samples to test a hypothesis that lagoons downstream from golf courses contain higher phosphate concentrations that those not downstream from golf courses. If:

A. you plan to use a two-sample t-test to compare phosphate concentrations, 
B. your significance level is 0.05,
C. you want to have a power of 0.80

...then how many samples would you have to collect if there is a SMALL (d ~ 0.2), MODERATE (d ~ 0.5), or LARGE (d ~ 0.8) effect size? First of all, how can we GUESS what the effect size might be if we have *a priori* information? It's *basically* the difference in means divided by the pooled standard deviation. So you could estimate what the effect size will be, or just try endpoints for low and high effect size.

To use pwr.t.test() function, there are four components:

n = sample size
d = Cohen's d effect size
sig.level = alpha
power = power (standard is ~ 0.8)

If you give the function THREE of those things, and set the fourth to NULL, then the fourth thing will be calculated for you.

```{r a_priori_power_calcs}

# Small effect size:
power_small <- pwr.t.test(n = NULL, d = 0.2, sig.level = 0.05, power = 0.8)
power_small # ~ 393 samples needed for each group

# Moderate effect size:
power_moderate <- pwr.t.test(n = NULL, d = 0.5, sig.level = 0.05, power = 0.8)
power_moderate # ~ 64 amples needed for EACH group

power_large <- pwr.t.test(n = NULL, d = 0.8, sig.level = 0.05, power = 0.8)
power_large # ~ 25 samples needed for EACH group

```

###3. Pipe to ggplot, highlight one column, interactivity in ggplotly

```{r}

graph_2 <- mortality %>% 
  filter(year == 2015, 
         ages == "All Ages", 
         sex == "Both Sexes", 
         race_and_hispanic_origin == "All Races-All Origins", 
         state != "United States") %>% 
  mutate(highlight = ifelse(state == "Kentucky", "Yes", "No")) %>%
  arrange(-death_rate) %>% 
  head(10) %>% 
    ggplot(aes(x = reorder(state, death_rate), y = death_rate)) +
    geom_col(aes(fill = highlight)) +
    labs(x = "", 
         y = "Drug-related death rates (2015)\n(deaths per 100,000 people)",
         title = "United States Drug-Related Mortality Rates",
         subtitle = "10 Highest Mortality Rates (2015)") +
    theme_classic() +
    scale_y_continuous(expand = c(0,0), 
                       limits = c(0,40), 
                       breaks = seq(0,40,by = 10)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), 
          axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 20)),
          legend.position = "none",
          text = element_text(family = "Times New Roman")) +
    scale_fill_manual(values = c("gray60","red"))

graph_2

# Use ggplotly to make it interactive
# Use tooltip to specify what you want to show up in the hovering menu (this can be customized, we'll just do simple customization here...)

ggplotly(graph_2, tooltip = "y")

```

Figure 1. Highest 10 drug-related mortality rates (deaths per 100,000 people) by state in 2015. Data from the Center for Disease Control [1].    

###4. Join (dplyr), annotation, labels, customization continued

We can use the ggrepel package (an add-on for ggplot) to add text labels to a graph and avoid overlapping labels. 

```{r, message = FALSE}

deaths_2015 <- mortality %>% 
    filter(year == 2015, 
         ages == "All Ages", 
         sex == "Both Sexes", 
         race_and_hispanic_origin == "All Races-All Origins")

# Make sure to look at the 'income' data frame

deaths_income <- full_join(deaths_2015, income) %>% 
  select(state, death_rate, med_income, population)

# Make a simple scatterplot of median income v. death rate
# Do with all states first, then show subsetting option within the ggplot line

death_income_scatter <- ggplot(filter(deaths_income, med_income > 60000), 
                               aes(x = med_income, 
                                   y = death_rate, 
                                   label = state)) +
  geom_point(aes(size = population, color = state), alpha = 0.7) +
  # geom_text() Try this first - show that they all overlap (centered over points)
  labs(x = "2016 Median Household Income (USD)", y = "Drug Related Death Rate (deaths per 100,000)") +
  geom_text_repel(size = 3, color = "gray50") +
  theme_classic() +
  theme(legend.position = "none") 
  

death_income_scatter

# ggplotly(death_income_scatter) # Note: ggrepel doesn't exist in plotly yet! 
```


###5. A finalized summary table (using kable and kableExtra)

```{r}

ca_table <- mortality %>% 
  filter(year >= 2010,
         state == "California", 
         ages == "All Ages", 
         sex == "Both Sexes", 
         race_and_hispanic_origin == "All Races-All Origins") %>% 
  select(year, deaths, population, death_rate)

# Make a table
# striped makes stripes (gray background rows)
# hover will highlight the row you're on 
# condensed makes row height smaller
ca_final <- kable(ca_table,
                  col.names = c("Year","Deaths","Population","Death Rate")) %>% 
   kable_styling(bootstrap_options = c("striped","hover"), 
                full_width = FALSE) %>% 
                row_spec(row = 4, bold = T, color = "magenta", background = "yellow") %>% 
  column_spec(column = 4, bold = T, color = "purple")

ca_final

```
