---
title: "penguins_anova"
author: "Allison Horst"
date: "November 1, 2018"
output: html_document
---

0. Start a new git-enabled Project by following along: 

- Fork my lab-6 repo in github
- Clone (copy URL)
- Open RStudio, and start a new project with Version Control using git
- Link to github by pasting URL when prompted
- Follow along by updating the markdown template
- Practice staging, committing and pushing back to github throughout lab

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Our question: "Do mean flipper lengths differ significantly for Adelie, Gentoo and Chinstrap penguins near Palmer Archipeligo?"

1. Load packages and read-in penguins.csv data
```{r}

library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")

```

2. Do some exploring (note that data is already in tidy format)

New functions: with + table for easy counts over multiple factors (this is a nice one...)
Update species information (simplify) using case_when

```{r}

counts <- with(penguins, table(species, sex))
View(counts)

# Update species information (common name) using case_when()
penguins2 <- penguins %>% 
  mutate(common_name = 
         case_when(
    species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
    species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
    species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
  )
  ) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")


```

Then...

- exploratory graphs
- summary statistics
- Levene's test (equal variances)
- one-way ANOVA (type III)
- use agricolae package to get groups

Exploratory graphs
```{r, include = FALSE}

hists <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(bins = 10) +
  facet_wrap(~ common_name, scale = "free")

hists

qqs <-  ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq(bins = 10) +
  facet_wrap(~ common_name, scale = "free")

qqs

# All groups approx. normally distributed. How do they look compared to each other?

box <-  ggplot(penguins2, aes(x = common_name, y = flipper_length)) +
  geom_boxplot(width = 0.2) +
  geom_jitter(width = 0.1, alpha = 0.5, aes(color = common_name))

box

```

Levene's test for equal variances (and keep in mind the general rule: if the greatest variances is < 4x bigger than the smallest variance, then usually those are "close enough" to assume equal variance)

```{r}
## DO LEVENE'S TEST HERE

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    mean = mean(flipper_length),
    sd = sd(flipper_length),
    variance = var(flipper_length)
  )

variances # Definitely close enough (also see that this lines up with Levene's test)

penguin_levene <- leveneTest(flipper_length ~ common_name, data = penguins2)

penguin_levene # Variances are not significantly different


```

One-Way ANOVA (Type III)

```{r}

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
penguin_sum <- summary(penguin_aov)

penguin_ph <- TukeyHSD(penguin_aov)

## Testing other options (for a variation that allows to call values using in-line references...) There's one aov.test() in the new 'onewaytests' package, BUT that doesn't have Tukey's for a post-hoc pairwise comparison...

penguin_factor <- data.frame(
  penguins2 %>% 
  mutate(name_factor = as.factor(common_name),
         flip_numeric = flipper_length
         )
  )

aov2 <- aov.test(flip_numeric ~ name_factor, data = penguin_factor)

# Yes, there are at least two means that differ significantly. 
# Our next question: which ones? 
# Bonferroni correction:

post_hoc <- paircomp(aov2, adjust.method = "bonferroni")

# All means are significantly different from each other! 


```

How do we report this outcome?

Mean flipper lengths for Adelie (mean ± sd, n = ), Gentoo (mean ± sd, n = ), and Chinstrap (mean ± sd, n =) penguins all differed significantly across all groups by one-way ANOVA (F(`r aov2$parameter`) = `r aov2$statistic`, *p* < 0.001, $\alpha$ = 0.05) with post-hoc pairwise comparison by Bonferroni correction (p < 0.001 for all pairwise comparisons).

