---
title: "Lab 9 - Multiple Linear Regression"
author: "Allison Horst"
date: "November 25, 2018"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Lab 9 Objectives:**

- Explore multivariate data (SLO housing prices)
- Perform multiple linear regression
- Assess diagnostics
- Compare different models by AIC
- Explain model outputs
- Make a nice table of regression results
- Make predictions using a final model

###1. Load packages

```{r packages, include = FALSE}

library(tidyverse)
library(corrplot)
library(stargazer)

```

###2. Load data, and filter to only include homes in SLO, Arroyo Grande, Santa Maria-Orcutt, and Atascadero

```{r}

homes <- read_csv("slo_homes.csv") # Read in data

homes_sub <- homes %>% 
  filter(City == "Arroyo Grande" | City == "San Luis Obispo" | City == "Atascadero" | City == "Santa Maria-Orcutt")

```

###3. Go exploring (visual) + think critically

*Note: It's OK to LOOK at things separately, even if you're including all in a model together!*

Example: if I want to compare distribution of housing prices by CITY (ignoring all other variables), I can do that.

```{r by_city}

mean_by_city <- homes_sub %>% 
  group_by(City) %>% 
  summarize(
    mean = mean(Price)
  )

by_city <- ggplot(homes_sub, aes(x = Price)) +
  geom_density(aes(color = City, fill = City), alpha = 0.3) + # Note: just to show what the geom_violin shows
  theme_classic() +
  scale_x_continuous(expand = c(0,0), limits = c(0,3e6)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = "Home Prices (USD)", y = "Density")

by_city

```

Or another question: Overall relationship between home square footage and price, separated by City? 

```{r by_sqft}

by_sqft <- ggplot(homes_sub, aes(x = SqFt, y = Price)) +
  geom_point(aes(color = City, pch = Status), alpha = 0.5) 

by_sqft

# Observations here: Does relationship appear ~ linear? Anything else we can pick out re: trends, outliers, etc.? What is the general trend? Any outliers? Is there reason enough for us to omit it?

```

```{r explore_data}

# Correlation matrix: any concerns about collinearity? 
# Histogram of final: 
# Relationships for all:

```

At this point: any major concerns? What is going to let us know if our assumptions are really violated? 

###4. Multiple linear regression

Multiple linear regression in R follows the same syntax we've been using so far: 

    lm(y ~ x1 + x2 + x3..., data = df_name)
    
Let's try this model a couple of different ways: 

(1) Use all available variables (saturated model) 
(2) Use only SqFt as a predictor for "home size" generally, and omit PricePerSqFt (since it's derived from two other existing variables in the model)

```{r saturated}

homes_lm1 <- lm(Price ~ City + Bedrooms + Bathrooms + SqFt + PricePerSqFt + Status, data = homes_sub) 

summary(homes_lm1)

# This makes no sense! Why not? Interpret several of the coefficients (for both continuous predictors and factor levels). What should we exclude? Include? Based on WHAT?

```

The next model: Excluding bedrooms and bathrooms, AND price per square foot...

```{r subset}

homes_lm2 <- lm(Price ~ City + SqFt + Status, data = homes_sub)

summary(homes_lm2) # NOW this is something that makes sense conceptually and mathematically...

# Interpret coefficients for City, SqFt, and Status. Do these make sense based on what you know about housing prices? 

```

Wait...but what if I wanted everything to be with respect to a Regular sale status? Then I need to change my factor levels. We've done this before, here we'll use a different function (fct_relevel) from *forcats* package in the tidyverse. 

```{r fct_relevel}

homes_sub$Status <- factor(homes_sub$Status)
# Check to ensure it's a factor now

class(homes_sub$Status) # Yay! A factor

# Check levels:
levels(homes_sub$Status) # Current reference level is 'Foreclosure'

# Reassign reference level of "Status" to "Regular":
homes_sub$Status <- fct_relevel(homes_sub$Status, "Regular")

# Now run the regression again - same equation, but now the reference levels are different! 
homes_lm3 <- lm(Price ~ City + SqFt + Status, data = homes_sub)

summary(homes_lm3) 

# Question: What happens if you run all of this, then go back to the code chunk above and run that code again? It'll change there TOO...remember, information in R is stored. 

# Now: How do we interpret the coefficients for Short Sale/Foreclosure vs. Regular sales? 

```

Interpret the statistical outcomes above. 

###5. Model diagnostics

Remember, since we're concerned about *residuals* (distance that actual observations exist from model predictions), we can only evaluate some assumptions *after* running the regression. 

Then we can evaluate model diagnostics using the plot() function:

```{r diagnostics}

plot(homes_lm3)

# Nothing really concerning...but there are errors. What variables do you think might be missing from the models that could account for some of the error?

# Some examples: Lot size (yard?), ocean view, etc.

# But overall, looks good and makes sense! 

```

###6. Model comparison by Akaike Information Criterion

The AIC is a quantitative metric for model "optimization" that balances complexity with model fit. The best models are the ones that fit the data as well as possible, as simply as possible. Recall: lower AIC value indicates a *more optimal* balance - **BUT STATISTICS IS NO SUBSTITUTE FOR JUDGEMENT!!!**

```{r AIC}

sat_aic <- AIC(homes_lm1) # 10699
final_aic <- AIC(homes_lm3) # 11148

# BUT WHICH ONE WOULD YOU PICK ANYWAY???? Only a ~4% difference in AIC value...not that different in the scheme of things, and the first one doesn't make sense...so still stick with the second one! 

```

###7. Regression tables with *stargazer*

```{r stargazer, results = 'asis'}
lm_tab <- stargazer(homes_lm1, homes_lm3, type = "html")
# Note: If you want to work with this in Word, save to html, open, copy and paste into Word. 
```

###8. Making predictions

Using your final selected model, predict the housing price for a range of home sizes, sale status, and city. 

The predict() function uses the following syntax:

      predict(model_name, newdata = new_data_name)
      
Defaults are to exclude the prediction SE and mean confidence interval - if you want to include, use arguments

      se.fit = TRUE
      interval = "confidence" 
      interval = "prediction"

First, you need to create a new data frame of values that contain ALL NECESSARY VARIABLES **with the same variable names AND level strings**.

```{r df_new}

# First, make a new data frame
# Note that the df_new created below has the SAME variable names and level strings as the original model data (otherwise R won't know how to use it...)
# Work through this on your own to figure out what it actually does:

df_new <- data.frame(City = rep(c("San Luis Obispo",
                                  "Santa Maria-Orcutt",
                                  "Atascadero",
                                  "Arroyo Grande"), 
                                each = 60), 
                     SqFt = rep(seq(from = 500,
                                    to = 3500, 
                                    length = 20), 
                                times = 12), 
                     Status = rep(c("Regular",
                                    "Foreclosure",
                                    "Short Sale"), 
                                  times = 12, 
                                  each = 20))

```

Make predictions for the new data using predict():

```{r predict}

price_predict <- predict(homes_lm3, newdata = df_new, se.fit = TRUE, interval = "confidence") # Makes prediction

# Bind to the data to make it actually useful:

predict_df <- data.frame(df_new, price_predict)

```

Then visualize it!

```{r graph, echo = FALSE, messages = "hide"}

ggplot(predict_df, aes(x = SqFt, y = fit.fit)) +
  geom_line(aes(color = City)) +
  geom_point(data = homes_sub, aes(x = SqFt, y = Price), alpha = 0.5) +
  facet_wrap(~Status) +
  labs(x = "Home Size (Sq. Ft.)", y = "Predicted Home Price ($)") +
  scale_x_continuous(limits = c(500,3500), breaks = seq(500, 3500, by = 1000)) +
  scale_y_continuous(limits = c(0,1.5e6))
  theme_light() 

```

END LAB